---
title: "Final Project"
author: "Michael Cheng and Ben Dreier"
date: "5/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(leaflet)
library(geojson)
```

## Intro

This Rmd file will present a few highlights of the full analysis present in the Final.R file.

First, let's load the data file. Let M be the data file.

```{r readin, echo=FALSE}
M <- read.csv("MembershipEdited.csv"); head(M)
```

## SECTION 1: Hometown analysis with Contingency Table

Is having a hometown that isn't Los Angeles correlated with being a Convert from another/no religion?
My hypothesis: In theory, it could make sense because people who moved from further away towns might have joined the church in search of community, then converted religions
While people originally from LA already had a community from growing up there; or another factor could dominate

First, let's find out the number of people from LA and the number of converts

```{r}

HomeLA <- which(M$Hometown == "Los Angeles, CA") #People with hometowns LA
length(HomeLA) #Los Angeles, CA has 1,232
HomeNotLA <- which(M$Hometown != "Los Angeles, CA") #All people who were not from LA 
length(HomeNotLA) #Not LA has 1,814
ConvertsA <-which(M$Former.Church == "Convert") #All the converts
length(ConvertsA) #427 converts
NotConverts <-which(M$Former.Church != "Convert") #All the non-converts
length(NotConverts) #2,619 non-converts

# And put it in a data fram
LA <- M$Hometown == "Los Angeles, CA"
Converts <- M$Former.Church == "Convert"
HomeR <- data.frame(LA, Converts) #make a dataframe with just the logical columns


```

We can run a chi squared test to see how dependent the variables of Being from LA and Being a Convert are:

```{r}

chisq.test(HomeR$LA,HomeR$Converts)

```

The p value is very small, so it is incredibly unlikely that the correlation between LA and Converts arose by chance. We can take this as evidence enough that there is a strong correlation between being from LA and being a Convert

```{r echo=FALSE}
slices <- c(424, 3, 808, 1818) #notice that the numbers are slightly different since I added back the residents who were excluded earlier
lbls <- c("Convert Locals", "Convert Migrants", "Non-Convert Locals", "Non-Convert Migrants")
pct <-  round(slices/sum(slices)*100, 2) 
lbls <- paste(lbls, slices, pct) # add percents to labels 
lbls <- paste(lbls,"%",sep="") # ad % to labels 
pie(slices,labels = lbls, col=rainbow(length(lbls)),
    main="Converts, Non-Converts, Migrants, and Locals at the Church")
```

## SECTION 3: Comparing Baptism with Distance Travelled using Permutation Test

Looking at the dataset, we see a lot of Baptist representation. Now, we want to figure out whether being a Baptist is correlated at all with Distance Traveled using a Permutation test.

The following code uses a word-search algorithm to guess which churches in the dataset are Baptist churches. 

```{r}

Churches <- (M$Former.Church) #column of all individuals churches
Denoms <- c("Baptist", "Methodist", "AME", "Episcopal", "CME", "Presbyterian", "Catholic") # Enumerate some words pointing to church denomination
# Function to get which churches contain which denomination strings
whichContain <- function(ch, dn){
  denoms <- rep("", length(ch))
  i <- 1
  for(c in ch){
    for(d in dn){
      #Check if the church has the denomination
      if( is.na(grepl(d, c)) ){break}
      if( grepl(d, c) ){
        denoms[i] = paste(denoms[i], d);
      }
    }
    i <- i + 1
  }
  return(denoms)
}
WhichDenoms <- whichContain(Churches, Denoms)
isBaptist <- WhichDenoms == " Baptist";  #true if an individual is Baptist

sum(isBaptist)

```

We found 1185 Baptist churches. Out of only 3,046 individuals, this represents a big part of the population.

Now, calculate the observed differences between Distance Travelled for Baptists vs Non-Baptists

```{r}
km <- M$Distance / 1000 #divide distance by 1000 to get distance in km
mean(km) #mean distance for all Individuals is 1580.57 km
median(km) #median distance for all Individuals is 2137 km

#Calculate the observed Distance Traveled for Baptists and non-Baptists
dBaptists <- sum((km) * (isBaptist == TRUE))/sum(isBaptist == TRUE) #average distance for Baptists is 2028.802 km
dOthers <- sum((km) * (isBaptist == FALSE))/sum(isBaptist == FALSE) #average distance for non-Baptists is 1295.159 km
observed <- dBaptists - dOthers; observed #on average, Baptists traveled 733.64 km farther than non-Baptists

```

Now we can perform a permutation test to determine if this difference could have arisen by chance.

```{r}
N <- 10000
diffs <- numeric(N)
for(i in 1:N){
  Samp <- sample(isBaptist); Samp #permuted isBaptist column
  dBaptists <- sum((km) * (Samp == TRUE))/sum(Samp == TRUE); dBaptists
  dOthers <- sum((km) * (Samp == FALSE))/sum(Samp == FALSE)
  diffs[i] = dBaptists - dOthers #as likely to be negative or positive
}

# The mean should hover near zero
mean(diffs)
```

We can plot a histogram of observed differences from the permutation test to visually confirm that the differences are normally distributed with a mean of 0

```{r echo=FALSE}
hist(diffs, breaks = "FD", xlim = c(-500,900), main = "Histogram of Observed Differences", xlab = "Observed Differences")
abline(v = observed, col = "red")
```

The red line, at the observed difference in the non-permuted data, is a statistical outlier. We can get its p-value:

```{r}
pvalue <- (sum(diffs >= observed)+1)/(N+1); pvalue
```

And see that there was about a 0.0001 chance of this difference occuring by chance. 

We can conclude that, for whatever reason, Baptists at the People's Independent Church migrated from further distances than non-Baptists, on average 733.64 km further.

## SECTION 6: Mapping the Hometown data

For visual analysis and presentation, we can create a map using the hometown addresses in the data set. Location data was generated in R using the Google Maps API, but now we load data from a saved backup instead of fetching it all again. We also have access to a json file defining where and how to plot state outlines.

```{r}
locs_file <- read.csv("locations2.csv")
states <- geojsonio::geojson_read("us-states.json", what="sp")
```

Now we count the number of times each state enumerated in the states object occurs in the dataset; this is what we will plot on our heatmap.

```{r}
counts = {}
for(i in 1:length(states$name)){
  curr_state <- toString(states$name[i])
  state_count <- length(M$State[M$State == curr_state])
  counts = c(counts, state_count)
}

# Put these counts in the states object used for visualization
states$count <- counts
```

After deciding how to delineate color bins and creating a palette, we can use leaflet to map how frequently each state appears.

```{r echo=FALSE}
states$count <- counts

bins <- c(0, 10^seq(0, 3.5, 0.5))
pal <- colorBin("YlOrRd", domain = states$count, bins = floor(bins))


m <- leaflet(states)
m <- addTiles(m)
m <- addPolygons(m, fillColor = ~pal(counts), weight=1, color="white", fillOpacity = 0.7)
m <- addLegend(m, "bottomright", pal=pal, values=counts)
m 
```

Then, using the individual location data gathered earlier, we plot every member on the map.

```{r echo=FALSE}
addCircleMarkers(m, lng=locs_file$lon, lat=locs_file$lat, label=M$Former.Church, radius=1)
```

Darker dots represent multiple individuals who live in the same town.